---
title: "EcoEncontros seminar series data"
author: "Melina Leite & Júlia Barreto"
date: "`r format(Sys.time(), '%d de %B de %Y')`"
output:
  rmdformats::readthedown:
    highlight: kate
    self_contained: true
    thumbnails: false
    lightbox: true
    gallery: false
  pdf_document:
    highlight: tango
    toc: yes
---
```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
library(tidyverse); library(cowplot); library(patchwork)
theme_set(theme_cowplot())
library(janitor); library(lubridate)

opts_chunk$set(fig.align = 'center', warning = FALSE, message = FALSE, error = FALSE, echo=T, cache=F)
options(formatR.arrow = TRUE, width = 90, help_type = "html")
```


# EcoEncontros Seminar talks

Talks from EcoEncontros Seminar series at the Graduate Program of Ecology in the University of São Paulo (PPGE-USP), Brazil

See file `metadata.txt`, in folder `data` for more description and detail of the dataset.
```{r}
data1 <- read.table("data/presentations_PPGE_2008-2019.csv", sep=",",
                   header=T, as.is=T)
data1$date <- dmy(data1$date)
data1$year <- year(data1$date) 
#skimr::skim(data)
```


**For ALL analyses:** Excluding special events as round tables and discussions not related to a project or study presented by someone.

```{r}
IDs <- c(154, 211, 212, 230, 244, 250, 289, 319)
data <- data1 %>% filter(!id %in% IDs)
```


# Population data from the graduate programm PPGE-USP from 2008-2019

Number of students, post-docs and professors in the PPGE-USP per gender and year.

```{r}
pop <- read.table("data/pop_PPGE_2008-2019.csv", sep=",",
                   header=T, as.is=T)
kable(pop)
```



# Data description

Descriptive totals: From `r dim(data1)[1]` talks in 12 years, we used `r dim(data)[1]`  as we excluded special seminars as round tables, presentations not related to the study project of the speaker. 

In total, `r paste0(round(((data %>% tabyl(gender))[2,3]*100)), "%")` of the talks were given by men (n=`r (data %>% tabyl(gender))[2,2]`) and `r paste0(round(((data %>% tabyl(gender))[1,3]*100)), "%")` by women (n=`r (data %>% tabyl(gender))[1,2]`).

```{r}
dim(data)
data %>% tabyl(gender)
```

# Speakers data

Proportion and number of male and female speakers per academic position.

Excluding talks from non-academic professionals.  N= `r dim(data)[1]`.

```{r}
data <- data %>% filter(position_cat != "others")
data$position_cat <- fct_relevel(data$position_cat, "student", 
                                 "postdoc","professor")
```


```{r, echo=F}
data %>% tabyl(position_cat, gender) %>% adorn_percentages("row") %>%
  adorn_pct_formatting(digits = 0) %>%
  adorn_ns() %>%
  kable(caption="Proportion and number (in parenthesis) of females (F) and males (M) per academic position category.")
```

```{r num speak}
n.y <- data %>% tabyl(position_cat, gender)
ggplot(data, aes(x=position_cat, fill=gender)) + geom_bar() + 
  ylab("Number of speakers") + xlab("") +
   scale_fill_manual("gender", values = c("#b2abd2", "#fdb863")) +
  scale_color_discrete(name="Gênero") +
  theme(text = element_text(size=18),
    axis.text.x = element_text(size=16)) +
  annotate("text", x=1:3, y=n.y$M+n.y$F/2, size=5,
           label = c("52%", "43%","24%"))

ggsave("figures/numberSpeakers_position.jpeg", units="in", width=7, height=4.5, dpi=300)
```

```{r num speak port}
n.y <- data %>% tabyl(position_cat, gender)
ggplot(data, aes(x=position_cat, fill=gender)) + geom_bar() + 
  ylab("Número de palestrantes") + xlab("") +
   scale_fill_manual("gênero", values = c("#b2abd2", "#fdb863")) +
  scale_x_discrete(labels = c("Pós-graduande", "Pós-doc", "Docente")) +
  scale_color_discrete(name="Gênero") +
  theme(text = element_text(size=18),
    axis.text.x = element_text(size=16)) +
  annotate("text", x=1:3, y=n.y$M+n.y$F/2, size=5,
           label = c("52%", "43%","24%"))

ggsave("figures/numberSpeakers_position_port.jpeg", units="in", width=7, height=4.5, dpi=300)
```

Variation in time.

```{r, echo=F}
prop.y <- data %>% tabyl(year,gender) %>%
  adorn_percentages("row") %>%
  adorn_pct_formatting(digits = 0) 
n.y <- data %>% tabyl(year,gender)

ggplot(data, aes(x=as.factor(year), fill=gender)) + geom_bar()  +
  ylab("Number of speakers") + xlab("") +
   scale_fill_manual("gender", values = c("#b2abd2", "#fdb863")) +
  theme(text=element_text(size=18),
    axis.text.x = element_text(size=16,angle = 45, hjust=1)) +
  annotate("text", x=1:12, y=n.y$M+n.y$F/2,size=5,
           label = prop.y$F)

ggsave("figures/numberSpeakers_year.jpeg", units="in", width=11, height=7, dpi=300)
```


### Origin of the speakers



```{r, echo=F}
origin <- paste(data$department, data$institute,data$university)
eco.ib <- sort(table(origin), decreasing = T)[1]
todo.ib <- length(grep("IB USP", origin, value=T))
```

There were `r sort(table(origin), decreasing = T)[1]` (`r round(eco.ib*100/dim(data)[1])`%) talks given by people from the PPGE population.

Including talks from the Institute of Biosciences, USP, there were `r todo.ib` (`r round(todo.ib*100/dim(data)[1])`%).

## PPGE-USP population data 

```{r echo=FALSE}
pop2 <- pop %>%  
  pivot_longer(2:7, names_to=c("category","gender"), values_to="N",
               names_sep = "_") %>%
  pivot_wider(names_from = "gender", values_from="N") %>%
  mutate(propFtotal = total_F/(total_F+total_M),
         ratioFtotal = total_F/total_M,
         propFcat = F/(F+M),
         ratioFcat = F/M) %>%
  dplyr::select(year, category, total_F, total_M, propFtotal,
                ratioFtotal,F,M, propFcat, ratioFcat)
```

PPGE-USP population size by gender in time

```{r, echo=F}
pop %>% dplyr::select(year,total_F,total_M) %>% pivot_longer(2:3, names_to = "gender", values_to = "N") %>%
ggplot(aes(x=as.factor(year), y=N, fill=gender)) + geom_col() +
   scale_fill_manual("gender", values = c("#b2abd2", "#fdb863"),
                     label=c("F", "M")) +
  xlab("")+ ylab("Population size")+
  theme(text=element_text(size=18),
    axis.text.x = element_text(size=16,angle = 45, hjust=1)) +
  annotate("text", x=1:12, y= pop$total_M + pop$total_F/2, size=5,
           label = paste0(round(pop$total_F*100/(pop$total_M+pop$total_F),0),"%"))
ggsave("figures/popSize_year.jpeg", units="in", width=11, height=7, dpi=300)
```

PPGE-USP population size by gender and position and year
```{r, fig.height=8}
pop2 %>% mutate(ytext = M + F/2) %>%
  pivot_longer(7:8, names_to = "gender", values_to = "N") %>%
  ggplot(aes(x=as.factor(year),y=N, fill=gender)) + geom_col() + 
  facet_wrap(~category, scales="free",ncol=1)+
  ylab("N") + xlab("") +
   scale_fill_manual("gender", values = c("#b2abd2", "#fdb863")) +
  theme(text = element_text(size=18),
    axis.text.x = element_text(size=16, angle=45, hjust=1)) +
  geom_text(aes(x=as.factor(year), y=ytext,
                label=paste0(round(propFcat*100), "%") ))

ggsave("figures/popSize_positionYear.jpeg", units="in", width=7, height=14, dpi=300)
```


## Comparing proportions of female speakers in the seminar and in the population by position.

```{r, fig.cap="Relationship between the proportion of females in the PPGE population in each category of academic position in each year (x axis) and the proportion of female speakers in the same category and year. Horizontal and vertical gray dashed lines are the 50% of each proportion and diagonal black dashed line indicates where the propotions are similar."}
propS <- data %>% group_by(year,position_cat,gender) %>% 
  count() %>% 
  pivot_wider(names_from = gender, values_from = n, values_fill = 0) %>%
  mutate(propFspeaker = F/(M+F)) 
proportions <- pop2 %>% dplyr::select(year, category, propFcat, F,M) %>% 
  rename("position_cat"="category") %>%
  left_join(propS, by=c("year", "position_cat")) %>%
  mutate(position_cat = fct_relevel(position_cat, "student", "postdoc", "professor"))
ggplot(proportions,aes(x=propFcat, y=propFspeaker, col=year)) +
  scale_color_gradient()+
  geom_point() +
  facet_wrap(~position_cat)+
  xlim(0,1) +
  xlab("Proportion of females in the population") +
  ylab("Proportion of females as speakers")+
  geom_abline(slope=1,intercept=0, linetype="dashed") +
  geom_hline(yintercept = 0.5, linetype="dashed", col="lightgray")+
  geom_vline(xintercept = 0.5, linetype="dashed", col="lightgray")
ggsave("figures/propFemale_popXspeaker.jpeg", units="in", width=14, height=5, dpi=300)
```


By year
```{r, fig.height=8}
proportions %>% pivot_longer(c(3,8), names_to = "data",values_to = "proportion") %>%
ggplot(aes(x=year, y=proportion, col=data)) +
  geom_point() +
  scale_color_manual(name="Dataset",values = 1:2,
                     labels=c("PPGE population", "Speakers pop"))+
  geom_smooth(method="lm") +
  facet_wrap(~position_cat, ncol=1)+
  ylab("Proportion of females")+
  geom_hline(yintercept = 0.5, linetype="dashed", col="lightgray")
ggsave("figures/propFemale_popXspeaker_byYear.jpeg", units="in", width=7, height=10, dpi=300)
```
