---
title: "Gender bias in speakers and career position"
author: "Melina Leite & Júlia Barreto"
date: "`r format(Sys.time(), '%d de %B de %Y')`"
output:
  rmdformats::readthedown:
    highlight: kate
    self_contained: true
    thumbnails: false
    lightbox: true
    gallery: false
  pdf_document:
    highlight: tango
    toc: yes
---
```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
library(tidyverse); library(cowplot); library(patchwork)
theme_set(theme_cowplot())
library(janitor); library(lubridate)
library(DHARMa); library(ggeffects)
library(bbmle); library(MASS)
library(hnp); library(performance)


opts_chunk$set(fig.align = 'center', warning = FALSE, message = FALSE, error = FALSE, echo=T, cache=F)
options(formatR.arrow = TRUE, width = 90, help_type = "html")
```

# Data

Data description and summary in script `0_data_summary`.

```{r}
data1 <- read.table("data/presentations_PPGE_2008-2019.csv", sep=",",
                   header=T, as.is=T)
data1$date <- dmy(data1$date)
data1$year <- year(data1$date) 
```

Excluding special events as round tables and discussions not related to a project or study presented by someone.

```{r}
IDs <- c(154, 211, 212, 230, 244, 250, 289, 319)
data <- data1 %>% filter(!id %in% IDs)
```

Excluding talks from non-academic professionals.

```{r}
data <- data %>% filter(position_cat != "others")
data$position_cat <- fct_relevel(data$position_cat, "student", 
                                 "postdoc","professor")
```

Creating dummy column to indicate if the speaker is a female (1) or not (0)
```{r}
data$fem <- 1
data$fem[data$gender == "M"] <- 0 
```

Including varible to indicated the affirmative actions period
```{r}
data$affirm_action <- ifelse(data$year<2018,"before", "after")
data$affirm_action <- fct_relevel(data$affirm_action,"before", "after")
```

There were `r table(data$affirm_action)[1]` seminars before and  `r table(data$affirm_action)[2]` after the affirmative actions.



Population data from PPGE-USP from 2008-2019, to create sex ratio per year and academic position.
```{r}
pop <- read.table("data/pop_PPGE_2008-2019.csv", sep=",",
                   header=T, as.is=T)
```



# Modeling

## Proportions of female speakers by academic position and time.

Binomial models with the response variable as 0 if the speaker is male or 1 if female. Response varibles as year, and academic position and affirmative actions.


```{r, echo=T}
mod0 <- glm(fem ~ 1, family=binomial, data= data)
mod1 <- glm(fem ~ year, family=binomial, data= data)
mod2 <- glm(fem ~ affirm_action, family=binomial, data= data)
mod3 <- glm(fem ~ position_cat, family=binomial, data= data)

mod4 <- glm(fem ~ position_cat + year, family=binomial, data= data)
mod5 <- glm(fem ~ position_cat*year, family=binomial, data= data)
mod6 <- glm(fem ~ position_cat + affirm_action, family=binomial, data= data)
mod7 <- glm(fem ~ position_cat*affirm_action, family=binomial, data= data)

kable(AICtab(mod0,mod1,mod2,mod3,mod4,mod5,mod6,mod7, base=T, weights=T), digits=2)

#as.data.frame(AICtab(mod0,mod1,mod2,mod3,mod4,mod5,mod6,mod7, base=T, weights=T)) %>%
#  mutate_at(c(1,2,4), round, digits=2) %>% kable()
```

```{r, echo=T}
library(lme4)
mod0 <- glmer(fem ~ 1 + ratioFcat + (1|year), family=binomial, data= data)
mod1 <- glmer(fem ~ scale(year) + ratioFcat+ (1|year), family=binomial, data= data)
mod2 <- glmer(fem ~ affirm_action + ratioFcat+ (1|year), family=binomial, data= data)
mod3 <- glmer(fem ~ position_cat + ratioFcat+ (1|year), family=binomial, data= data)

mod4 <- glmer(fem ~ position_cat + ratioFcat+ scale(year) + (1|year), 
              family=binomial, data= data)
mod5 <- glmer(fem ~ position_cat*scale(year)+ ratioFcat + (1|year), 
              family=binomial, data= data)
mod6 <- glmer(fem ~ position_cat + affirm_action + ratioFcat + (1|year), 
              family=binomial, data= data)
mod7 <- glmer(fem ~ position_cat*affirm_action + ratioFcat + (1|year), 
              family=binomial, data= data)

kable(AICtab(mod0,mod1,mod2,mod3,mod4,mod5,mod6,mod7, base=T, weights=T), digits=2)

#as.data.frame(AICtab(mod0,mod1,mod2,mod3,mod4,mod5,mod6,mod7, base=T, weights=T)) %>%
#  mutate_at(c(1,2,4), round, digits=2) %>% kable()
```



Best model is the one with interaction between academic position and affirmative actions (as time event), but it is equaly plausible with the model with year as time event and only academic position.

## Residual diagnostic of the selected models

All equaly plausible models presented satisfactory residual diagnostic.
```{r}
hnp(mod7)
hnp(mod5)
hnp(mod3)
```

```{r}
plot(simulateResiduals(mod7))
plot(simulateResiduals(mod5))
plot(simulateResiduals(mod3))
```


## Models results

```{r}
summary(mod7)
performance::r2_nagelkerke(mod7)
my7 <- ggpredict(mod7, terms=c("position_cat","affirm_action"))
plot(my7) +
  geom_hline(yintercept = 0.5, linetype="dashed")
```
The best fitted model includes only academic position and affirmative actions and the interaction between them (R2 = `r round(performance::r2_nagelkerke(mod7),2)`, Figure 1).

Figure proportion of female speakers by position_cat e affirm_action


```{r fig prop}
suma <- data %>% count(position_cat, affirm_action,fem)
prs <- as.data.frame(my7)

ggplot(suma, aes(x=position_cat, y=fem,col=affirm_action))+
  geom_point(aes(, size=n),position=position_dodge(0.6), alpha=0.2, show_guides=F) +
  scale_size(range=c(1,10), breaks = c(3,10,20,60))+
  geom_pointrange(data=prs, aes(x=x, y=predicted, col=group,
                                ymax=conf.high, ymin=conf.low), 
             position=position_dodge(0.6)) +
  geom_hline(yintercept = 0.5, linetype="dashed", col='gray') +
  scale_color_manual(name="Affirmative \n actions", 
                     values = c("goldenrod", "green4")) +
  ylab("Proportion of female speakers")+
  xlab("Academic position")+
  theme(text = element_text(size=20),
        axis.text = element_text(size=18))
ggsave("figures/prop_female_speakers.jpeg", width=9, height = 6)  
```


```{r fig prop port}
suma <- data %>% count(position_cat, affirm_action,fem)
prs <- as.data.frame(my7)

ggplot(suma, aes(x=position_cat, y=fem,col=affirm_action))+
  geom_point(aes(, size=n),position=position_dodge(0.6), alpha=0.2, show_guides=F) +
  scale_size(range=c(1,10), breaks = c(3,10,20,60))+
  geom_pointrange(data=prs, aes(x=x, y=predicted, col=group,
                                ymax=conf.high, ymin=conf.low), 
             position=position_dodge(0.6)) +
  geom_hline(yintercept = 0.5, linetype="dashed", col='gray') +
  scale_color_manual(name="Ações \n afirmativas", 
                     label = c("antes", "depois"),
                     values = c("goldenrod", "green4")) +
  scale_x_discrete(labels=c("Pós-graduande", "Pós-doc", "Docente"))+
  ylab("Proporção de mulheres palestrantes")+
  xlab("")+
  theme(text = element_text(size=20),
        axis.text = element_text(size=18))
ggsave("figures/prop_female_speakers_port.jpeg", width=9, height = 6)  
```


```{r}
summary(mod5)
performance::r2(mod5)
my5 <- ggpredict(mod5, terms=c("year", "position_cat"))
plot(my5) +
  geom_hline(yintercept = 0.5, linetype="dashed")
```

```{r}
summary(mod3)
performance::r2(mod3)
my3 <- ggpredict(mod3, terms=c("position_cat"))
plot(my3) +
  geom_hline(yintercept = 0.5, linetype="dashed")
```


## Proportions of female speakers by academic position and time - controling by population proportions

The models below controls by differences in gender ratios in PPGE population by academic position. For that, we included a control variable in every model in the model set with the ratio of females in the year and by position.
```{r}
data <- data %>% left_join(pop2[,c(1:2,10)], by=c("year", "position_cat"="category"))
```


```{r, echo=T}
modc0 <- glm(fem ~ 1 + ratioFcat, family=binomial, data= data)
modc1 <- glm(fem ~ year + ratioFcat, family=binomial, data= data)
modc2 <- glm(fem ~ affirm_action + ratioFcat, family=binomial, data= data)
modc3 <- glm(fem ~ position_cat+ ratioFcat, family=binomial, data= data)

modc4 <- glm(fem ~ position_cat + year + ratioFcat, family=binomial, 
             data= data)
modc5 <- glm(fem ~ position_cat*year + ratioFcat, family=binomial, data= data)
modc6 <- glm(fem ~ position_cat +affirm_action + ratioFcat, family=binomial,
             data= data)
modc7 <- glm(fem ~ position_cat*affirm_action + ratioFcat, family=binomial,
             data= data)

AICtab(modc0,modc1,modc2,modc3,modc4,modc5,modc6,modc7, base=T, weights=T)
```

Similar results as without the ratio of females in the PPGE population

## Residual diagnostic of the selected models

All equaly plausible models presented satisfactory residual diagnostic.
```{r}
par(mfrow=c(2,2))
hnp(modc7)
hnp(modc5)
hnp(modc3)
```

```{r}
plot(simulateResiduals(modc7))
plot(simulateResiduals(modc5))
plot(simulateResiduals(modc3))   
```

## Models results

Using 1:1 population gender ratio
```{r}
summary(modc7)
performance::r2(modc7)
myc7 <- ggpredict(modc7, terms=c("position_cat","affirm_action",
                                 "ratioFcat[1]"))
plot(myc7) +
  geom_hline(yintercept = 0.5, linetype="dashed")
```


```{r fig prop}
suma <- data %>% count(position_cat, affirm_action,fem) %>%
  mutate(position_cat = fct_relevel(position_cat, "student", "postdoc", "professor"))
prs <- as.data.frame(myc7)

ggplot(suma, aes(x=position_cat, y=fem,col=affirm_action))+
  geom_point(aes(, size=n),position=position_dodge(0.6), alpha=0.2, show_guides=F) +
  scale_size(range=c(1,10), breaks = c(3,10,20,60))+
  geom_pointrange(data=prs, aes(x=x, y=predicted, col=group,
                                ymax=conf.high, ymin=conf.low), 
             position=position_dodge(0.6)) +
  geom_hline(yintercept = 0.5, linetype="dashed", col='gray') +
  scale_color_manual(name="Affirmative \n actions", 
                     values = c("goldenrod", "green4")) +
  ylab("Proportion of female speakers")+
  xlab("Academic position")+
  theme(text = element_text(size=20),
        axis.text = element_text(size=18))
ggsave("figures/prop_female_speakers_model.jpeg", width=9, height = 6)  
```

```{r}
summary(modc5)
performance::r2(modc5)
myc5 <- ggpredict(modc5, terms=c("year", "position_cat",
                                 "ratioFcat[1]"))
plot(myc5) +
  geom_hline(yintercept = 0.5, linetype="dashed")
```

```{r}
summary(modc3)
performance::r2(modc3)
myc3 <- ggpredict(modc3, terms=c("position_cat", "ratioFcat[1]"))
plot(myc3) +
  geom_hline(yintercept = 0.5, linetype="dashed")
```

<!-- # Efficiency of affirmataive actions for professors -->

<!-- ```{r} -->
<!-- data.prof <- data %>% filter(position_cat == "professor") -->
<!-- ``` -->

<!-- Model before affirmative action (2018) -->
<!-- ```{r} -->
<!-- data.prof.before <- data.prof %>% filter(affirm_action =="before") -->
<!-- data.prof.after <- data.prof %>% filter(affirm_action =="after") -->

<!-- mbefore <- glm(fem ~ year, family=binomial, data= data.prof.before) -->
<!-- mbefore.null <- glm(fem ~ 1, family=binomial, data= data.prof.before) -->
<!-- AICctab(mbefore,mbefore.null) -->

<!-- mafter <- glm(fem ~ year, family=binomial, data= data.prof.after) -->
<!-- mafter.null <- glm(fem ~ 1, family=binomial, data= data.prof.after) -->
<!-- AICctab(mafter,mafter.null) -->

<!-- summary(mbefore) -->
<!-- summary(mafter) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- nulo <- data.frame(intercept = rep(c(arm::invlogit(coef(mbefore.null)), -->
<!--               arm::invlogit(coef(mafter.null))), c(10,2)), -->
<!--             year=rep(2008:2019), -->
<!--             affirm_action=rep(c("before", "after"), c(10,2))) -->

<!-- ggplot(nulo, aes(y=intercept,x=year,col=affirm_action))+ geom_line() + -->
<!--     ylim(0,0.8) + -->
<!--       geom_hline(yintercept = 0.5, linetype="dashed", col="gray")+ -->
<!-- ``` -->


<!-- ```{r} -->
<!-- mbef <- ggpredict(mbefore, terms=c("year")) -->
<!-- plot(mbef) + -->
<!--   geom_hline(yintercept = 0.5, linetype="dashed") -->

<!-- maft <- ggpredict(mafter, terms=c("year")) -->
<!-- plot(maft) + -->
<!--   geom_hline(yintercept = 0.5, linetype="dashed") -->


<!-- bind_rows(list(before = as.data.frame(mbef), after=as.data.frame(maft)),  -->
<!--           .id="time") %>% -->
<!--   ggplot(aes(x=x, y=predicted, col=time))+ geom_line() + -->
<!--   geom_hline(yintercept = 0.5, linetype="dashed", col="gray")+ -->
<!--   geom_ribbon(aes(ymin=conf.low, ymax=conf.high), alpha=0.3, col="white")+ -->
<!--   ylim(0,0.8) -->

<!-- ``` -->

