---
title: "Gender bias in speakers and career position"
author: "Melina Leite & Júlia Barreto"
date: "`r format(Sys.time(), '%d de %B de %Y')`"
output:
  rmdformats::readthedown:
    highlight: kate
    self_contained: true
    thumbnails: false
    lightbox: true
    gallery: false
  pdf_document:
    highlight: tango
    toc: yes
---
```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
library(tidyverse); library(cowplot); library(patchwork)
theme_set(theme_cowplot())
library(DHARMa); library(ggeffects)
library(bbmle);
library(performance)
library(glmmTMB)

opts_chunk$set(fig.align = 'center', warning = FALSE, message = FALSE, error = FALSE, echo=T, cache=F)
options(formatR.arrow = TRUE, width = 90, help_type = "html")
```

# Data

Data description and summary in script `0_data_summary`.

```{r}
load("data_clean/data_proportion.Rdata")
load("data_clean/data_pop_ppge.Rdata")
data <- data_p
```


Creating dummy column to indicate if the speaker is a female (1) or not (0)
```{r}
data$fem <- 1
data$fem[data$gender == "M"] <- 0 
```


# Modeling

Proportions of female speakers by academic position, before and after affirmative actions. Controling for population proportions: the ratio of females at the graduate program by academic position and year.

```{r}
data <- data %>% left_join(pop2[,c(1:2,10)], by=c("year", "position_cat"="category")) %>%
  mutate(position_cat = fct_relevel(position_cat, "student","postdoc",
                                    "professor"))
```

Mixed effects model, including year as random intercept.

Binomial distribution

```{r, echo=T}
mod0 <- glmmTMB(fem ~ 1 + ratioFcat + (1|year), 
              family=binomial, data= data)
mod1 <- glmmTMB(fem ~ affirm_action + ratioFcat + (1|year), 
              family=binomial, data= data)
mod2 <- glmmTMB(fem ~ position_cat + ratioFcat + (1|year), 
              family=binomial, data= data)
mod3 <- glmmTMB(fem ~ position_cat + affirm_action + ratioFcat + (1|year), 
              family=binomial, data= data)
mod4 <- glmmTMB(fem ~ position_cat*affirm_action + ratioFcat + (1|year), 
              family=binomial, data= data)

kable(AICtab(mod0,mod1,mod2,mod3,mod4, base=T, weights=T), digits=2)
```


## Residual diagnostic of the selected models

```{r}
plot(simulateResiduals(mod4))
plot(simulateResiduals(mod2))
```

## Models results

Using 1:1 population gender ratio
```{r}
summary(mod4)
performance::r2(mod4)
my4 <- ggpredict(mod4, terms=c("position_cat","affirm_action",
                                 "ratioFcat[1]"))
plot(my4) +
  geom_hline(yintercept = 0.5, linetype="dashed")
```


```{r fig prop}
suma <- data %>% count(position_cat, affirm_action,fem) %>%
  mutate(position_cat = fct_relevel(position_cat, "student", "postdoc", "professor"))
prs <- as.data.frame(my4)

ggplot(suma, aes(x=position_cat, y=fem,col=affirm_action))+
  geom_point(aes(, size=n),position=position_dodge(0.6), alpha=0.2, show_guides=F) +
  scale_size(range=c(1,10), breaks = c(3,10,20,60))+
  geom_pointrange(data=prs, aes(x=x, y=predicted, col=group,
                                ymax=conf.high, ymin=conf.low), 
             position=position_dodge(0.6)) +
  geom_hline(yintercept = 0.5, linetype="dashed", col='gray') +
  scale_color_manual(name="Affirmative \n actions", 
                     values = c("goldenrod", "green4")) +
  ylab("Proportion of female speakers")+
  xlab("Academic position")+
  theme(text = element_text(size=20),
        axis.text = element_text(size=18))
ggsave("figures/prop_female_speakers_model.jpeg", width=9, height = 6)  
```

```{r}
summary(mod2)
performance::r2(mod2)
my2 <- ggpredict(mod2, terms=c( "position_cat",
                                 "ratioFcat[1]"))
plot(my2) +
  geom_hline(yintercept = 0.5, linetype="dashed")
```

