---
title: "Gender bias in audience of seminars and career position"
author: "Melina Leite, Júlia Barreto & Isabella Romitelli"
date: "`r format(Sys.time(), '%d de %B de %Y')`"
output:
  rmdformats::readthedown:
    highlight: kate
    self_contained: true
    thumbnails: false
    lightbox: true
    gallery: false
  pdf_document:
    highlight: tango
    toc: yes
---
```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
library(tidyverse); library(cowplot); library(patchwork)
theme_set(theme_cowplot())
library(janitor); library(lubridate)
library(DHARMa); library(ggeffects)
library(bbmle); library(MASS)
library(hnp); library(performance)
library(FactoMineR); library(factoextra)


opts_chunk$set(fig.align = 'center', warning = FALSE, message = FALSE, error = FALSE, echo=T, cache=F)
options(formatR.arrow = TRUE, width = 90, help_type = "html")
```

# Data

Data description and summary in script `0_data_summary`.


```{r}
load("data_clean/data_audience.Rdata")
data <- data_a
```


# Modeling


Negative binomial

Tests glmm
```{r}
library(glmmTMB)
mg0 <- glmmTMB(audience_n~ 1 + (1|year), data=data, family= nbinom2)
mg1 <- glmmTMB(audience_n~ gender + (1|year), data=data, family= nbinom2)
mg2 <- glmmTMB(audience_n~ position_cat + (1|year), data=data, family= nbinom2)
mg3 <- glmmTMB(audience_n~ affirm_action + (1|year), data=data, family= nbinom2)

mg4 <- glmmTMB(audience_n~ gender + position_cat+ (1|year), 
               data=data, family= nbinom2)
mg5 <- glmmTMB(audience_n~ gender + affirm_action+ (1|year), 
               data=data, family= nbinom2)
mg6 <- glmmTMB(audience_n~ affirm_action + position_cat+ (1|year), 
               data=data, family= nbinom2)

mg7 <- glmmTMB(audience_n~ gender * position_cat+ (1|year), 
               data=data, family= nbinom2)
mg8 <- glmmTMB(audience_n~ gender * affirm_action+ (1|year), 
               data=data, family= nbinom2)
mg9 <- glmmTMB(audience_n~ affirm_action * position_cat+ (1|year), 
               data=data, family= nbinom2)

mg10 <- glmmTMB(audience_n~ gender + position_cat + affirm_action + (1|year), 
               data=data, family= nbinom2)
mg11 <- glmmTMB(audience_n~ gender * position_cat + affirm_action + (1|year), 
               data=data, family= nbinom2)
mg12 <- glmmTMB(audience_n~ gender + position_cat * affirm_action + (1|year), 
               data=data, family= nbinom2)

mg13 <- glmmTMB(audience_n~ gender * position_cat * affirm_action + (1|year), 
               data=data, family= nbinom2)

AICtab(mg0,mg1,mg2, mg3, mg4,mg5,mg6,mg7,mg8,mg9,mg10,mg11,mg12,mg13,      base=T, weights=T) %>% kable(digits=2)
```

```{r}
summary(mg11)
```








```{r echo=TRUE}
mg0 <- glm.nb(audience_n~ 1, data=data)
mg1 <- glm.nb(audience_n~ gender, data=data)
mg2 <- glm.nb(audience_n~ position_cat, data=data)
mg3 <- glm.nb(audience_n~ year, data=data)
mg3b <- glm.nb(audience_n~ affirm_action, data=data)

mg4 <- glm.nb(audience_n~ gender + position_cat, data=data)
mg5 <- glm.nb(audience_n~ gender + year, data=data)
mg5b <- glm.nb(audience_n~ gender + affirm_action, data=data)
mg6 <- glm.nb(audience_n~ year + position_cat, data=data)
mg6b <- glm.nb(audience_n~ affirm_action + position_cat, data=data)

mg7 <- glm.nb(audience_n~ gender*position_cat, data=data)
mg8 <- glm.nb(audience_n~ gender*year, data=data)
mg8b <- glm.nb(audience_n~ gender*affirm_action, data=data)
mg9 <- glm.nb(audience_n~ year*position_cat, data=data)
mg9b <- glm.nb(audience_n~ affirm_action*position_cat, data=data)

mg10 <- glm.nb(audience_n~ gender + position_cat + year, data=data)
mg10b <- glm.nb(audience_n~ gender + position_cat + affirm_action, data=data)
mg11 <- glm.nb(audience_n~ gender*position_cat + year, data=data)
mg11b <- glm.nb(audience_n~ gender*position_cat + affirm_action, data=data)
mg12 <- glm.nb(audience_n~ gender + position_cat * year, data=data)
mg12b <- glm.nb(audience_n~ gender + position_cat * affirm_action, data=data)
mg13 <- glm.nb(audience_n~ gender*year + position_cat, data=data)
mg13b <- glm.nb(audience_n~ gender*affirm_action + position_cat, data=data)


mg14 <- glm.nb(audience_n~ gender*position_cat*year, data=data)
mg14b <- glm.nb(audience_n~ gender*position_cat*affirm_action, data=data)

AICtab(mg2, mg0,mg1, mg3, mg4,mg5,mg6,mg7,mg8,mg9,mg10,mg11,mg12,mg13,mg14,
      mg3b,mg5b,mg6b,mg8b,mg9b,mg10b,mg11b,mg12b,mg13b, mg14b,
      base=T, weights=T) %>% kable(digits=2)
```


## Residual diagnostic

```{r}
hnp::hnp(mg11b)
plot(simulateResiduals(mg11b))
```

```{r}
hnp::hnp(mg10b)
plot(simulateResiduals(mg10b))
```

## Models result

The two equally plausible models for the audience included gender, academic position and affirmative actions as predictors, with the difference that the best fitted model includes an interaction of gender and academic position (R2 = `r round(performance::r2(mg11b)$R2_Nagelkerke,2)`)

```{r echo=TRUE}
summary(mg11b)
performance::r2(mg11b)
```

Average audience before affirmative actions:`r exp(coef(mg11b)[1])`

Average audience aftter affirmative actions: `r exp(coef(mg11b)[1]+ coef(mg11b)[5])`


```{r}
myg11b <- ggpredict(mg11b, terms=c("position_cat","gender", "affirm_action"))
as.data.frame(myg11b)%>% mutate(predicted = round(predicted,digits=0))
```


Complete figure
```{r}
prs <- as.data.frame(myg11b) %>% rename(affirm_action = facet)
colnames(prs)[1] <- "position_cat"
kable(prs , digits = 0)
```


```{r}
ggplot(data, aes(x=position_cat, y=audience_n)) +
  geom_point(aes(col=gender), position = position_dodge(0.6), alpha=0.3,
             size=3,show.legend = F) +
  facet_grid(~affirm_action, 
             labeller = as_labeller(c("before"="Before affirmative actions" , "after"="After affirmative actions" ))) +
 #scale_color_manual(values = c("#b2abd2", "#fdb863")) +
  scale_color_manual(values = c("#6D57CF","#FCA532")) +
  scale_fill_manual(name="Gender", values = c("#6D57CF","#FCA532")) +
  geom_pointrange(data=prs, aes(x=position_cat, y=predicted,fill=group,
                                ymax=conf.high, ymin=conf.low), alpha=1,
             position=position_dodge(0.6), size=1, shape=21, col="black") +
  xlab("Academic position") + ylab("Audience (N)") 

ggsave("figures/audience_speakers.jpeg", width=8, height = 4)  
```


```{r}
ggplot(data, aes(x=affirm_action, y=audience_n)) +
  geom_point(aes(col=gender), position = position_dodge(0.6), alpha=0.3,
             size=3,show.legend = F) +
  facet_grid(~position_cat, labeller = 
               as_labeller(c(student = "Pós-graduande",
                             postdoc =  "Pós-doc",
                             professor = "Docente"))) +
  scale_color_manual(values = c("#6D57CF","#FCA532")) +
  scale_x_discrete(labels = c("Antes", "Depois"))+
  scale_fill_manual(name="Gênero", values = c("#6D57CF","#FCA532")) +
  geom_pointrange(data=prs, aes(x=affirm_action, y=predicted,fill=group,
                                ymax=conf.high, ymin=conf.low), alpha=1,
             position=position_dodge(0.6), size=1.2, shape=21, col="black") +
    xlab("Ações afirmativas") + ylab("Audiência (N)") 


ggsave("figures/audience_speakers_b_port.jpeg", width=8, height = 4)  
```


figura apresentacao IB-mulheres
```{r}
ggplot(data=prs, aes(x=position_cat, y=predicted,fill=group,
                                ymax=conf.high, ymin=conf.low)) +
  geom_pointrange(, alpha=1,
             position=position_dodge(0.6), size=1.2, shape=21, col="black") +
  facet_grid(~affirm_action, labeller= as_labeller(c(before = "Antes",
                                                   after = "Depois"))) +
  scale_color_manual(values = c("#6D57CF","#FCA532")) +
  scale_x_discrete(labels =  c("Pós-graduande", "Pós-doc", "Docente"))+
  scale_fill_manual(name="Gênero", values = c("#6D57CF","#FCA532")) +
    xlab("") + ylab("Audiência (N)") 

ggsave("figures/audience_speakers_C_port.jpeg", width=8, height = 4)  
```

```{r}
prs %>% filter(position_cat == "professor", affirm_action == "after") %>%
ggplot(aes(x=position_cat, y=predicted,fill=group,
                                ymax=conf.high, ymin=conf.low)) +
  geom_pointrange(alpha=1,
             position=position_dodge(0.6), size=1.2, shape=21, col="black") +
  scale_color_manual(values = c("#6D57CF","#FCA532")) +
  scale_x_discrete(labels =  c("Docente"))+
  scale_fill_manual(name="Gênero", values = c("#6D57CF","#FCA532")) +
    xlab("") + ylab("Audiência (N)") 

ggsave("figures/audience_speakers_D_port.jpeg", width=4, height = 4)  
```


```{r echo=TRUE}
summary(mg10b)
performance::r2(mg10b)
```
```{r}
myg10b <- ggpredict(mg10b, terms=c("position_cat","gender", "affirm_action"))
plot(myg10b) +
  scale_color_manual(values = c("#b2abd2", "#fdb863"))
```




# Only professors - productivity metrics

Investigating if differences in productivity between male and female professors and researches are related to the audience.

Measured productivity publication metrics from Google Scholar for professors and researchers. 

Creating productivity index using PCA 1st axis from metrics.

## PCA productivity metrics

```{r}
dp <- data %>% filter(!is.na(data$total_citation_n),
                      !is.na(data$nature_index_count))
table(dp$gender, dp$affirm_action)
```

Productivity publication metrics

```{r}
pca1 <- PCA(dp[, c(22:29)], graph=F)
p1 <- fviz_pca_biplot(pca1, col.ind = dp$gender, addEllipses=TRUE,
                      col.ind.sub="none",  geom="point",
                      repel = TRUE) +
  geom_vline(xintercept = 0, linetype="dashed") + 
  geom_hline(yintercept = 0, linetype="dashed")+
  scale_color_manual(name="Gênero",values = c("#6D57CF","#FCA532"))+
  scale_shape(name="Gênero")+
  scale_fill_manual(name="Gênero",values = c("#6D57CF","#FCA532"))+
  ggtitle("Métricas desempenho professores") +
  xlab("PC1 (52%)") + ylab("PC2 (21%)") +
  theme_cowplot()

p1
ggsave("figures/pca_biplot.jpeg", width=6, height = 6)  
```
For the analysis specific for professor talks (N=`r length(dp$id)`), the PCA results show that all the productivity metrics for professors were highly correlated (Figure 2B) with the first axis (52%of variance explained) while the institution indexes composed the second PCA axis (21% of variation explained).

Extracting PCA 2 first axes
```{r}
dp$pc1 <- pca1$ind$coord[,1]
dp$pc2 <- pca1$ind$coord[,2]
```

# Modeling

We used generalized linear models with negative binomial distribution, given the large variation in the audience (range between `r paste0(range(data$audience_n)[1], " and ", range(data$audience_n)[2])`).
 
```{r}
m0 <- glm.nb(audience_n ~ 1+affirm_action, data=dp)
m1 <- glm.nb(audience_n ~ gender+affirm_action, data=dp)
m2 <- glm.nb(audience_n ~ affirm_action +pc1, data=dp)

m3 <- glm.nb(audience_n ~ affirm_action +gender + pc1 , data=dp)

m4 <- glm.nb(audience_n ~ affirm_action +gender*pc1, data=dp)

AICtab(m0,m1,m2,m3,m4,
       base=T, weights=T) %>% kable(digits=2)
```


```{r}
m0 <- glmmTMB(audience_n ~ 1 + affirm_action +(1|year), data=dp, family=nbinom2)
m1 <- glmmTMB(audience_n ~ gender +affirm_action +(1|year), data=dp, family=nbinom2)
m2 <- glmmTMB(audience_n ~ pc1 + pc2 +affirm_action +(1|year), data=dp, family=nbinom2)

m3 <- glmmTMB(audience_n ~ gender + pc1 + pc2 +affirm_action +(1|year), 
              data=dp, family=nbinom2)

m4 <- glmmTMB(audience_n ~ gender*pc1 + gender*pc2 +affirm_action +(1|year),
              data=dp, family=nbinom2)

AICtab(m0,m1,m2,m3,m4,
       base=T, weights=T) %>% kable(digits=2)
#summary(m3)
```


## Residual diagnostic

Best model
```{r}
hnp(m3)
plot(simulateResiduals(m3))
```

## Model results

```{r echo=TRUE}
summary(m3)
performance::r2(m3)
```

We used the first and second PCA axis as predictors together with gender to explain the professor's audience, and found that, as expected, audience increases with productivity index (first PCA axis) but female professors still presented on average audience *1.3* times smaller than male professors (R2 of the best fitting model = `r performance::r2(m3)`).


```{r}
my3 <- ggpredict(m3, terms=c("pc1","gender"))
my3 <- as.data.frame(my3)
ggplot(my3, aes(x=x, y=predicted, col=group)) +
  geom_ribbon(aes(ymin=conf.low,ymax=conf.high, fill=group), alpha=0.3,
             colour = NA) +
    geom_line()+
  scale_color_manual(name="Gender",values = c("#6D57CF","#FCA532"))+
  scale_fill_manual(name="Gender",values = c("#6D57CF","#FCA532"))+
   theme_cowplot() + ggtitle("") +
  ylab("Audience (N)") + xlab("Productivity index (PC1 axis)")+
  geom_point(data=dp, aes(x=pc1, y=audience_n, col=gender), alpha=0.6)
ggsave("figures/audience_professor.jpeg", width=9, height = 6)  
```

```{r}
my3 <- ggpredict(m3, terms=c("gender"))
plot(my3)
```

PC 2 - nao importante, nao "significativo"":

```{r}
my3 <- ggpredict(m3, terms=c("pc2","gender"))
plot(my3) +
  scale_color_manual(values = c("#b2abd2", "#fdb863"))+
  scale_fill_manual(values = c("#b2abd2", "#fdb863"))+
   theme_cowplot()
```


# Figure audience

```{r}
prs <- as.data.frame(myg11b) %>% rename(affirm_action = facet,
                                        position_cat=x)
f1<- ggplot(data, aes(x=affirm_action, y=audience_n)) +
  geom_point(aes(col=gender), position = position_dodge(0.6), alpha=0.3,
             size=3,show.legend = F) +
  facet_grid(~position_cat) +
 #scale_color_manual(values = c("#b2abd2", "#fdb863")) +
  scale_color_manual(values = c("#6D57CF","#FCA532")) +
  scale_fill_manual(name="Gender", values = c("#6D57CF","#FCA532")) +
  geom_pointrange(data=prs, aes(x=affirm_action, y=predicted,fill=group,
                                ymax=conf.high, ymin=conf.low), alpha=1,
             position=position_dodge(0.6), size=1, shape=21, col="black") +
  ylab("Audience (N)")  +
  xlab("Affirmative actions")+
  labs(tag="A")
  


my3 <- ggpredict(m3, terms=c("pc1","gender"))
my3 <- as.data.frame(my3)
my3$prof <- "Professors only"
f2 <- ggplot(my3, aes(x=x, y=predicted, col=group)) +
  geom_ribbon(aes(ymin=conf.low,ymax=conf.high, fill=group), alpha=0.3,
             colour = NA) +
    geom_line(size=1.5)+
  facet_grid(~prof)+
  scale_color_manual(name="Gender",values = c("#6D57CF","#FCA532"))+
  scale_fill_manual(name="Gender",values = c("#6D57CF","#FCA532"))+
   theme_cowplot() + ggtitle("") +
  ylab("Audience (N)") + xlab("Productivity index (PC1 axis)")+
  geom_point(data=dp, aes(x=pc1, y=audience_n, col=gender), alpha=0.6,
             size=2)+
  theme(legend.position="none") +
  labs(tag="C")


f3<- plot_spacer()

design <- "
           1111
           #22#
"

f1 + f2  + 
  plot_layout( design=design, guides="collect")
ggsave("figures/FIG_audience.jpeg", width=9, height = 8)  
```


```{r}

p1<- fviz_pca_biplot(pca1, col.ind = dp$gender, addEllipses=TRUE,
                      col.ind.sub="none",  geom="point",
                      repel = TRUE) +
  facet_grid(.~.)+
  geom_vline(xintercept = 0, linetype="dashed") + 
  geom_hline(yintercept = 0, linetype="dashed")+
  scale_color_manual(name="gender",values = c("#6D57CF","#FCA532"))+
  scale_shape(name="gender")+
  scale_fill_manual(name="gender",values = c("#6D57CF","#FCA532"))+
  labs(title="Professors' productivity", tag="B") +
  xlab("PC1 (52%)") + ylab("PC2 (21%)") +
  theme_cowplot() +
  theme(legend.position="none",
        plot.title = element_text(size=12, vjust=-5, hjust=0)) #+
  #coord_cartesian(clip = "off")+
  # scale_x_continuous(limits=c(-6,8), expand=c(0,0))+
  # scale_y_continuous(limits=c(-4,8), breaks=c(-4,-2,0,2,4,6))
  # annotate("rect", xmin=-6, xmax=8, ymin=7,ymax=8, fill="gray85")+
  #annotate("text",label="Professors'productivity", x=0, y=7.5, fill="gray85")



f1/(p1+f2) +plot_layout(guides="collect")
ggsave("figures/FIG_audience_test.jpeg", width=9, height = 8)  
```











