---
title: "Text analysis: title and abstract of male and female speakers"
author: "Melina Leite & Júlia Barreto"
date: "`r format(Sys.time(), '%d de %B de %Y')`"
output:
  rmdformats::readthedown:
    highlight: kate
    self_contained: true
    thumbnails: false
    lightbox: true
    gallery: false
  pdf_document:
    highlight: tango
    toc: yes
---
```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
library(tidyverse); library(cowplot); library(patchwork);library(ggbeeswarm)
theme_set(theme_cowplot())
library(janitor)
library(tidytext)
library(scales); library(ggrepel)
library(lubridate)
library(tidytext)
library(topicmodels)
library(tm); 
library(quanteda); library(quanteda.textplots)

opts_chunk$set(fig.align = 'center', warning = FALSE, message = FALSE, error = FALSE, echo=T, cache=F)
options(formatR.arrow = TRUE, width = 90, help_type = "html")
```

# Abstracts

```{r}
data <- read.table("data/presentations_PPGE_2008-2019.csv", sep=",",
                   header=T, as.is=T)
data$date <- dmy(data$date)
data$year <- year(data$date) 
#skimr::skim(data)
```

Excluding special events as round tables and discussions not related to a project or study presented by someone.
```{r}
IDs <- c(154, 250, 211, 289, 230, 167)
data <- data %>% filter(!id %in% IDs)
```

Using abstracts in English (original or translated)
```{r}
data <- data  %>% filter(!is.na(abstract_english)) 
```

Number of abstracts per group
```{r}
table(data$gender)
table(data$position_cat,data$gender)
```


Tidytext
```{r}
text_tok <- data %>% dplyr::select(id,gender,position_cat, audience_n,
                             abstract_english, title_english) %>%
  mutate(text = paste(title_english, abstract_english)) %>%
  unnest_tokens(output=word,input=text)

stop_w <- tibble(word = stopwords("en"))

# remove stopwords
text <- text_tok %>% 
  anti_join(stop_w, by="word")  %>% arrange(word) 

# remove numbers and other characters
text <- text %>% slice(-c(1:290)) %>% # number and some symbols
          filter(nchar(word)!=1) %>% # letters alone
          filter(!word %in% c("mpas", "δ13c", "β") )# remove acronyms, symbols


# solving some simple plurals
plural <- c("actions","advances", "adaptations", "amphibians", "animals",
            "ants","anurans","abundances","adjustments","adults","affects",
            "applications","approaches", "bees","builds", "birds","palms",
            "cerrados","challenges", "outputs",  "queens", "techniques",
            "continents","crops", "consequences", "questions",
            "decisions","declines","determines","determinants", "defenses",
            "dynamics",
            "economics", "ecosystems","environments", "experiences",
            "forests","grasslands",
            "genetics","gifts","gradients","guides","impacts",
            "increases","interactions","lives",
            "landscapes","males","mammals", "mangroves","models","movements",
            "mutualisms","networks","neotropics",
            "opilions","phenotypes","plants","projects","paths", "perspectives",
            "populations","promotes","relationships", "relations",
            "resources","responses","roads","services","skulls","snakes","seeds",
            "spaces", "spiders","stages", "trees", "variations",
            "threats")

text$word[text$word %in% plural] <- 
  substr(text$word[text$word %in% plural],
       1,nchar(text$word[text$word %in% plural])-1)
```


- Grouping similar words:

```{r}
lemma <- rbind(c("adaptive", "adaptation"),
               c("abilities","ability"),
               c("advancement", "advance"),
                  c("absent","absence"),
               c("agricultural", "agriculture"),
               c("agro", "agriculture" ),
               c("amazonia","amazon" ),
               c("amazonian","amazon" ),
               c("andean","andes"),
               c("apply","application"),
               c("applying","application"),
               c("apidae","apis"),
               c("arachnida","arachnid"),
               c("argue","argument"),
               c("basal", "basis"),
               c("behavioral","behavior"),
               c("behavioural","behavior"),
               c("bignonieae", "bignoniaceae"),
               c("biological", "biology"),
               c("brazilian","brazil"),
               c("building","build"),
               c("changing", "change"),
               c("cnidarian", "cnidaria"),
               c("coastal","coast"),
               c("colour", "color"),
               c("colors", "color"),
               c("communities","community" ),
               c("competitive", "competition"),
               c("complexity", "complex"),
               c("convergences", "convergence"),
               c("convergent", "convergence"),
               c("cordatus","cordata" ),
               c("croplands","crop"),
               c( "cultural", "culture"),
               c("darwin's", "darwin"),
               c("darwinian", "darwin"),
               c("defensive", "defense"),
               c("dependent","dependence"),
               c("detecting","detection"),
               c("determine", "determinant"),
               c("developmental", "development"),
               c("dispersers","dispersal"),
               c("disturbed", "disturbance"),
               c("diversification", "diversity"),
               c("dragonflies", "dragonfly"),
               c("drier", "drought"),
               c("ecological", "ecology"),
               c("ecologists", "ecology"),
               c("endemic", "endemism"),
               c("effectiveness", "efficiency"),
               c("environmental", "environment"),
               c("evolutionary", "evolution"),
               c("expanding", "expansion"),
               c("extinct", "extinction"),
               c("facilitate", "facilitation"),
               c("fisheries", "fishery"),
               c("floral", "flora"),
               c("floristic", "flora"),
               c("forested", "forest"),
               c("functional", "function"),
               c("functionally", "function"),
               c("functioning", "function"),
               c("frequencies", "frequency"),
               c("frequently", "frequency"),
               c("frequent", "frequency"),
               c("geographical", "geographic"),
               c("heterogeneties", "heterogeneity"),
               c("heterogeneous", "heterogeneity"),
               c("histories", "history"),
               c("integrated", "integration"),
               c("intregating", "integration"),
               c("integrative", "integration"),
               c("invasive", "invasion"),
               c("isotopic", "isotope"),
               c("linking", "link"),
               c("living", "live"),
               c("mammalia", "mammal"),
               c("managed", "manage"),
               c("managers", "manage"),
               c("mathematical", "mathematics"),
               c("mates", "mating"),
               c("mediated", "mediate"),
               c("mechanistic", "mechanism"),
               c("matrices", "matrix"),
               c("migratory", "migration"),
               c("mimicking", "mimicry"),
               c("modeling", "model"),
               c("mutualistic", "mutualism"),
               c("natural", "nature"),
               c("neotropical", "neotropic"),
               c("northeastern", "northeast"),
               c("occuring", "occur"),
               c("onça", "onca"),
               c("opiliones", "opilion"),
               c("parasite", "parasitism"),
               c("parent", "parenting"),
               c("phylogenies", "phylogeny"),
               c("phylogenetic", "phylogeny"),
               c("phylogenomic", "phylogeny"),
               c("pollinators", "pollination"),
               c("protected", "protect"),
               c("protective", "protect"),
               c("rainfall", "rain"),
               c("reconstructing", "reconstruction"),
               c("regulatory", "regulation"),
               c("regulates", "regulation"),
               c("relation", "relationship"),
               c("reproductive", "reproduction"),
               c("restored", "restoration"),
               c("robustness", "robust"),
               c("scientific", "science"),
               c("scientist", "science"),
               c("sexy", "sexual"),
               c("simulated", "simulation"),
               c("societies", "society"),
               c("social", "society"),
               c("socio", "society"),
               c("space", "spatial"),
               c("spacio", "spatial"),
               c("stabilize", "stability"),
               c("stable", "stability"),
               c("stories", "story"),
               c("strategic", "strategy"),
               c("strategies", "strategy"),
               c("structured", "structure"),
               c("structuring", "structure"),
               c("studies", "study"),
               c("studing", "study"),
               c("sustainable", "sustainability"),
               c("theories", "theory"),
               c("theoretical", "theory"),
               c("threatened", "threat"),
               c("tropical", "tropic"),
               c("vision", "visual")
               )
lemma <- as.data.frame(lemma)

for (i in 1:dim(lemma)[1]){
  text$word[text$word == lemma[i,1]] <- lemma[i,2]
}
```


# WORDS - all data

```{r}
table(text$gender)
table(text$position_cat ,text$gender)
```

Mean number of words by title+abstract
```{r}
text %>% count(id,gender) %>%
  ggplot(aes(x=gender, y=n)) + 
  geom_violin() + geom_boxplot(width=0.2)+
  ylab("Number of words in title + abtract")
  ggbeeswarm::geom_quasirandom(size=3, shape=21) 
```



20 more common workds
```{r}
text %>%
  count(word, sort = TRUE) %>% 
  top_n(20,n)%>%
  kable()
```

## Word cloud

```{r}
textplot_wordcloud(x=dfm(tokens(text$word)))
```

```{r}
par(mfrow=c(1,2))
textplot_wordcloud(x=dfm(tokens(text$word[text$gender=="F"])),
                   col="#6D57CF")
par(new=T)
textplot_wordcloud(x=dfm(tokens(text$word[text$gender=="M"])),
                   col="#FCA532")
```


## Word frequencies by gender

```{r}
props <- text %>%
  count(gender, word) %>%
  group_by(gender) %>%
  mutate(proportion = n / sum(n)) %>% 
  pivot_wider(names_from = gender, values_from = c(proportion,n)) %>%
  mutate(abs.dif.p = abs(proportion_F-proportion_M),
         rel.dif.p = pmax(proportion_F, proportion_M)/
           pmin(proportion_F, proportion_M)) %>%
  arrange(desc(abs.dif.p))
props$label <- NA
props$label[1:20] <- props$word[1:20]
```


```{r}
ggplot(props, aes(x=proportion_M,, y=proportion_F,
       color=abs.dif.p)) + 
  geom_abline(color = "gray40", lty = 2) +
  #geom_point(size=2.5, alpha=0.5)+
  geom_jitter(size=2.5, alpha=0.2)+
  geom_text_repel(aes(label=label), size=3.2)+
  scale_x_log10(name="Male most used words",
                labels = percent_format()) +
  scale_y_log10(name="Female most used words",
                labels = percent_format()) +
  scale_color_gradient(name="Abs Diff",low = "blue", high = "red",
                       labels=percent_format())  +
  theme(legend.justification = c(1, -0.1), legend.position = c(1, 0))
 # geom_smooth(method="lm")
ggsave("figures/abstract_wordFreq.jpg", height = 5, width=7)
```

Words that are close to the dashed line have similar frequencies in both genders. Words that are far from the line are words that are found more in one set of texts than another.

Legend: absolute differences in the frequency of the word by males and females. Differences above 0.3% are also indicated in text.

Correlation of word frequeency use between gender:
```{r}
cor.test(props$proportion_F, props$proportion_M)
```

Highly correlated -> it means they tend to use the same frequency of main word

20 words with the largest differences in frequency

```{r}
prop2 <- props %>% filter(!is.na(label)) %>%
  arrange(desc(proportion_F), desc(proportion_M)) %>%
  mutate(ntot = n_F + n_M) %>%
  mutate(word = fct_reorder(word,(ntot),max),
         proportion_F = proportion_F*-1) %>%
  pivot_longer(2:3,names_to = "gender", values_to ="proportion")

ggplot(prop2, aes(x=proportion, y=word,fill=gender)) +
  geom_col()+ ylab("") + xlab("Proportion")+
  scale_fill_manual(name="gender", values=c("#6D57CF","#FCA532"),
                    labels=c("F", "M"))+
   geom_vline(xintercept = c(-0.02,-0.01,0,0.01,0.02),
              linetype="dotted",
             col="darkgray") +
  scale_x_continuous(breaks=c(-0.02,-0.01,0,0.01,0.02),
                     labels = c(0.02,0.01,0,0.01,0.02))
  
ggsave("figures/abstract_wordFreq_barplot.jpeg", units="in", width=7, height=7, dpi=300)
```

## TF IDF

```{r}
text_id <- text %>% count(gender, word) %>% 
  bind_tf_idf(word, gender, n) %>%
  arrange(desc(tf_idf))
```

10 "exclusive" words for each group
```{r}
text_id$word <- as.factor(text_id$word)
text_id %>%
  group_by(gender) %>% 
  arrange(desc(tf_idf)) %>% 
  top_n(10, tf_idf) %>%  
  ggplot(aes(x = tf_idf, y = reorder(word, tf_idf), fill = gender)) +
  geom_col(show.legend = FALSE) +
  labs(x = NULL, y = "tf-idf") +
  facet_wrap(~gender, scales = "free") +
  theme_minimal()
```



# WORDS - professors only data

```{r}
textP <- text %>% filter(position_cat == "professor")

table(textP$gender)
```

Mean number of words by abstract
```{r}
textP %>% count(id,gender) %>%
  ggplot(aes(x=gender, y=n)) + 
  geom_violin() + geom_boxplot(width=0.2)+
  ggbeeswarm::geom_quasirandom(size=3, shape=21) 
```

20 most commmon words
```{r}
textP %>%
  count(word, sort = TRUE) %>% 
  top_n(20,n)%>%
  kable()
```

## Words Frequency by gender

```{r}
propsP <- textP %>%
    count(gender, word) %>%
  group_by(gender) %>%
  mutate(proportion = n / sum(n)) %>% 
  pivot_wider(names_from = gender, values_from = c(proportion,n)) %>%
  mutate(abs.dif.p = abs(proportion_F-proportion_M),
         rel.dif.p = pmax(proportion_F, proportion_M)/
           pmin(proportion_F, proportion_M)) %>%
  arrange(desc(abs.dif.p))
propsP$label <- NA
propsP$label[1:20] <- propsP$word[1:20]
```

```{r}
ggplot(propsP, aes(x=proportion_M, y=proportion_F,
       color=abs.dif.p)) + 
  geom_abline(color = "gray40", lty = 2) +
 # geom_point(size=2.5, alpha=0.3) +
  geom_jitter(size=2.5, alpha=0.3)+
  geom_text_repel(aes(label=label), size=3)+
  scale_x_log10(name="Male most used words",   limits=c(0.0003,0.02),
                labels = percent_format()) +
  scale_y_log10(name="Female Most used words", limits=c(0.0003,0.02),
                labels = percent_format()) +
  scale_color_gradient(name="Abs Diff",low = "blue", high = "red",
                       labels=percent_format())  +
  theme(legend.justification = c(1, -0.1), legend.position = c(1, 0))
 # geom_smooth(method="lm")
ggsave("figures/abstract_wordFreq_Prof.jpg", height = 5, width=7)
```

Words that are close to the dashed  line in these plots have similar frequencies in both genders. Words that are far from the line are words that are found more in one set of texts than another.

Legend: absolute differences in the frequency of the word by males and females.

Labels for the 20 words with largest differences in frequency.

Correlation of word frequeency use between gender:
```{r}
cor.test(propsP$proportion_F, propsP$proportion_M)
```


20 words with the largest differences in frequency

```{r}
propP2 <- propsP %>% filter(!is.na(label)) %>%
  arrange(desc(proportion_F), desc(proportion_M)) %>%
  mutate(ntot = n_F + n_M) %>%
  mutate(word = fct_reorder(word,(ntot),max),
         proportion_F = proportion_F*-1) %>%
  pivot_longer(2:3,names_to = "gender", values_to ="proportion")

ggplot(propP2, aes(x=proportion, y=word,fill=gender)) +
  geom_col()+ ylab("") + xlab("Proportion")+
  scale_fill_manual(name="gender", values=c("#6D57CF","#FCA532"),
                    labels=c("F", "M"))+
   geom_vline(xintercept = c(-0.02,-0.01,0,0.01,0.02),
              linetype="dotted",
             col="darkgray") +
  scale_x_continuous(breaks=c(-0.02,-0.01,0,0.01,0.02),
                     labels = c(0.02,0.01,0,0.01,0.02))
  
ggsave("figures/abstract_wordFreq_barplot_Prof.jpeg", units="in", width=7, height=7, dpi=300)
```


## TF IDF

```{r}
text_idP <- textP %>% count(gender, word) %>% 
  bind_tf_idf(word, gender, n) %>%
  arrange(desc(tf_idf))
```

10 "exclusive" words for each group
```{r}
text_idP$word <- as.factor(text_idP$word)
text_idP %>%
  group_by(gender) %>% 
  arrange(desc(tf_idf)) %>% 
  top_n(10, tf_idf) %>%  
  ggplot(aes(x = tf_idf, y = reorder(word, tf_idf), fill = gender)) +
  geom_col(show.legend = FALSE) +
  labs(x = NULL, y = "tf-idf") +
  facet_wrap(~gender, scales = "free") +
  theme_minimal()
```

# Topic model - all data

```{r}
matext <- text %>% count(id, gender, word) %>% mutate(id = paste(id, gender, sep="_")) %>%
  select(-gender) %>%
                cast_dtm(term=word,document=id,value=n)
```

Choosing number of topics: comparing AIC

```{r}
ap_lda2 <- LDA(matext, k = 2, control = list(seed = 1234))
ap_lda3 <- LDA(matext, k = 3, control = list(seed = 1234))
ap_lda4 <- LDA(matext, k = 4, control = list(seed = 1234))
ap_lda5 <- LDA(matext, k = 5, control = list(seed = 1234))
bbmle::AICtab(ap_lda2, ap_lda3, ap_lda4,ap_lda5, base=T)
```

two-topics model seems the most plausible model

## Word-topic probabilities

10 words with the largest probabilities for each group
```{r}
ap_topics <- tidy(ap_lda2, matrix = "beta")
ap_top_terms <- ap_topics %>%
      group_by(topic) %>%
      top_n(10, beta) %>%
      ungroup() %>%
      arrange(topic, -beta)
ap_top_terms %>%
mutate(term = reorder(term, beta)) %>% ggplot(aes(term, beta, fill = factor(topic))) + geom_col(show.legend = FALSE) +
facet_wrap(~ topic, scales = "free") + coord_flip()
```



## Document-topic probabilities - classifying the abstracts

and comparing the two groups by gender (if there is a difference in frequency)

```{r}
ap_documents <- tidy(ap_lda2, matrix = "gamma")
classifi <- ap_documents %>% mutate(gender = substr(document, nchar(document), nchar(document))) %>%
  group_by(document,gender) %>%
      top_n(1, gamma) 

table(classifi$gender, classifi$topic)

classifi %>% tabyl(gender, topic) %>% adorn_percentages() %>%
  adorn_pct_formatting(digits = 0) %>%
  adorn_ns() %>% kable()
```


```{r}
classifi %>%
 # mutate(title = reorder(title, gamma * topic)) %>%
  ggplot(aes(as.character(topic), gamma)) +
  geom_boxplot() +
  facet_wrap(~ gender)
```

## Chi-square test 
```{r}
chisq.test(classifi$gender, classifi$topic)
```


# Topic model - Professors only

```{r}
matextP <- textP %>% 
  count(id, gender, word) %>% mutate(id = paste(id, gender, sep="_")) %>%
  select(-gender) %>%
                cast_dtm(term=word,document=id,value=n)
```

```{r}
ap_lda2P <- LDA(matextP, k = 2, control = list(seed = 1234))
ap_lda3P <- LDA(matextP, k = 3, control = list(seed = 1234))
ap_lda4P <- LDA(matextP, k = 4, control = list(seed = 1234))
bbmle::AICtab(ap_lda2P, ap_lda3P, ap_lda4P,base=T)
```

## word-topic probabilities
```{r}
ap_topicsP <- tidy(ap_lda2P, matrix = "beta")
ap_top_termsP <- ap_topicsP %>%
      group_by(topic) %>%
      top_n(10, beta) %>%
      ungroup() %>%
      arrange(topic, -beta)
ap_top_termsP %>%
mutate(term = reorder(term, beta)) %>% ggplot(aes(term, beta, fill = factor(topic))) + geom_col(show.legend = FALSE) +
facet_wrap(~ topic, scales = "free") + coord_flip()
```

## Document-topic probabilities

```{r}
ap_documentsP <- tidy(ap_lda2P, matrix = "gamma")
classifiP <- ap_documentsP %>% mutate(gender = substr(document, nchar(document), nchar(document))) %>%
  group_by(document,gender) %>%
      top_n(1, gamma) 

table(classifiP$gender, classifiP$topic)

library(janitor)
classifiP %>% tabyl(gender, topic) %>% adorn_percentages() %>% 
  adorn_pct_formatting(digits = 0) %>%
  adorn_ns() %>% kable()
```


```{r}
classifiP %>%
 # mutate(title = reorder(title, gamma * topic)) %>%
  ggplot(aes(as.character(topic), gamma)) +
  geom_boxplot() +
  geom_violin()+
  facet_wrap(~ gender)
```

Chi-square test 
```{r}
chisq.test(classifiP$gender, classifiP$topic)
```


# Sentiment analysis

Chapter 2, Silge & RObinson. 2018

- The NRC lexicon categorizes words in a binary fashion (“yes”/“no”) into categories of positive, negative, anger, anticipation, disgust, fear, joy, sadness, surprise, and trust.
```{r}
get_sentiments("nrc")
```

- The Bing lexicon categorizes words in a binary fashion into positive and negative categories.
```{r}
get_sentiments("bing")
```

- The AFINN lexicon assigns words with a score that runs between -5 and 5, with neg‐ ative scores indicating negative sentiment and positive scores indicating positive sen‐ timent.
```{r}
get_sentiments("afinn")
```

**PENSAR**: tem que levar em conta número de palavras diferentes entre abstracts - principalmente se ouver diferença média de número de palavras por abstract de homens e mulehres né? ou não? 

## Score words difference in female and male abstracts

### All data

```{r}
affword <- get_sentiments("afinn")

affc <- text %>%
  count(id,gender,word, sort = TRUE) %>%
  inner_join(affword, "word")
```

Calculating the mean of the scores for each abtract (weighted by number of time the word appears) by gender:

```{r}
affc %>% group_by(id, gender) %>%
  summarise(mean.score = mean(value),
            weig.score = weighted.mean(value,n)) %>%
  ggplot(aes(x=gender,y=weig.score)) +
  geom_violin() +
  geom_boxplot(width=0.1) +
    geom_quasirandom()+
  ggtitle("Mean words score per abstract and gender")
```

## Professors

```{r}
affword <- get_sentiments("afinn")

affcP <- textP %>%
  count(id,gender,word, sort = TRUE) %>%
  inner_join(affword, "word")
```

Calculating the mean of the scores for each abtract (weighted by number of time the word appears) by gender:

```{r}
affcP %>% group_by(id, gender) %>%
  summarise(mean.score = mean(value),
            weig.score = weighted.mean(value,n)) %>%
  ggplot(aes(x=gender,y=weig.score)) +
  geom_violin() +
  geom_boxplot(width=0.1) +
    geom_quasirandom()+
  ggtitle("Mean words score per abstract and gender")
```


## Frequency of sentiment words per abstract

As classificações das palavras não me parecem muito acuradas com a linguagem científica. 

Precisa saber como ponderar pelo total de palavras.

### All data
```{r, fig.height=8}
nrcword <- get_sentiments("nrc")

nrc <- text %>%
  count(id,gender,word, sort = TRUE) %>%
  inner_join(nrcword, "word") %>%
  group_by(id,gender,sentiment) %>%
  summarise(n= sum(n))

ggplot(nrc, aes(x=gender, y=n)) +
  facet_wrap(~sentiment) +
  geom_violin() +
    geom_quasirandom()
```

### Professors

```{r, fig.height=8}
nrcword <- get_sentiments("nrc")

nrc <- textP %>%
  count(id,gender,word, sort = TRUE) %>%
  inner_join(nrcword, "word") %>%
  group_by(id,gender,sentiment) %>%
  summarise(n= sum(n))

ggplot(nrc, aes(x=gender, y=n)) +
  facet_wrap(~sentiment) +
  geom_violin()+
    geom_quasirandom()
```

```{r}
nrc %>% filter(sentiment == "positive") %>%
ggplot( aes(x=gender, y=n)) +
  geom_violin() +
  geom_boxplot(width=0.2) +
    geom_quasirandom()+
  ggtitle("Positive words")
```


## Frequency of sentiment words per abstract

## All data
```{r}
bingword <- get_sentiments("bing")

bing <- text %>%
  count(id,gender,word, sort = TRUE) %>%
  inner_join(bingword, "word") %>%
  group_by(id,gender,sentiment) %>%
  summarise(n= sum(n))

ggplot(bing, aes(x=sentiment, y=n)) +
  facet_wrap(~gender) +
  geom_violin()+
    geom_quasirandom()
```

## Professors

```{r}
bingword <- get_sentiments("bing")

bing <- textP %>%
  count(id,gender,word, sort = TRUE) %>%
  inner_join(bingword, "word") %>%
  group_by(id,gender,sentiment) %>%
  summarise(n= sum(n))

ggplot(bing, aes(x=sentiment, y=n)) +
  facet_wrap(~gender) +
  geom_violin() +
  geom_boxplot(width=0.2) +
    geom_quasirandom()
```



